### 2016-02-01 paper.tex
Initial ShareLaTeX Import

\documentclass[12pt]{article}

% sorting=none will order citations in order of appearance
% use sorting=nyt to sort by name, year, title
\usepackage[numbers, backend=biber,autolang=hyphen, bibencoding=inputenc, strict, isbn=false, uniquename=false, sorting=none, sortcites=true, style=numeric, autocite=superscript, url=false]{biblatex} % biblatex setup
\DeclareLanguageMapping{american}{american-apa}
\appto{\bibfont}{\footnotesize} % for 10pt bibliography
%\usepackage{xunicode}
\usepackage[american]{babel}
\usepackage[babel]{csquotes}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows,shapes,fit}
\usepackage{standalone}
\usepackage{mdframed}
\usepackage{capt-of}
\usepackage{changepage}
\usepackage{titlesec}
\usepackage[parfill]{parskip}

\titleformat*{\section}{\normalsize\bfseries} % Makes section titles 12 pt font
\titleformat{\section}{\normalsize\bfseries}{}{}{}[]
\titleformat{\subsection}{\bfseries}{}{}{}[]

\let\textcite=\autocite
\addbibresource{../../computing.bib}

\title{\large \textbf{Leveraging Historical Ties between Cognitive Science and
  Computer Science to Guide Innovative Use of Computers in Education}}

% Author info isn't included for the Annual Conference but some regional conferences might request it.
\author{}
%\author{\normalsize Author Name\\
%\normalsize email@example.com\\
%\normalsize Name of Your Department\\\
%\normalsize Your Institution Name}
\date{}

\makeatletter % This gets the margins for the title set.
\patchcmd{\@maketitle}{\begin{center}}{\begin{adjustwidth}{0.5in}{0.5in}\begin{center}}{}{}
\patchcmd{\@maketitle}{\end{center}}{\end{center}\end{adjustwidth}}{}{}
\makeatother

\begin{document}
\raggedright
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

% conflict between teaching programming syntax, teaching ``how to program'', teaching high-level concepts like abstractions and how to ``do'' problem solving.
\section{Coding for Everyone}
In the past few years, there has been increasing interest in a
movement to get more people, regardless of background, to learn to
program, in fact President Obama recently made a statement calling on
all children to have the opportunity to learn about computer
science\autocite{whitehouse_computer_2016}. Sites such as code.org
promote CS education opportunities for all citing statistics about
STEM jobs and arguing that in the 21st century knowledge about
computer science is foundational\autocite{code.org_every_2016}. Many
institutions of higher education either have programming requirements
for non-computer-science majors, or have been expanding programming to
non-majors\autocite{rich_cs1_2004,forte_motivation_2005,guzdial_design_2005}. At
the authors of this paper's institution, all first year general
engineering students are required to complete two 6-week rotations of
an introductory programming curriculum as part of their first-year
engineering course.

Juxtaposed with the observation that technology is changing rapidly,
and the half-life of technical skills is decreasing (cite half-life
article) it would seem teaching programming is a questionable use of
class time: the language and paradigm students are taught will likely
not be the same they are asked to use in their first job
post-graduation. However, if we acknowledge that learning to program
is more than simply learning to write code, and in particular if we
draw on work from the early era of computing which closely linked
programming to problem solving and cognitive skills, we can argue that
regardless of language or paradigm, learning to program may actually
be a way to learn to problem solve, reason about and organize
knowledge, and conceptualize processes.

While there are some recent studies which investigate programming
education as a means of teaching problem-solving, there is also a
large body of research that primarily focuses on teaching students the
details of how to code. We argue that in most cases these details are
less important than centering programming education around
problem-solving skills and if educators hope to achieve this goal then
all aspects of introductory programming courses, from the course
objectives, to the curriculum used, to the assessments administered,
must be chosen intentionally with this goal in mind.

%We posit that as long as researchers focus on only one aspect of a
%programming course, whether it be the language to use, the paradigm to
%teach, the assessment method to use, etc., there will continue to be a
%lack of con consensus of what approach is best, just as
%\textcite{pears_survey_2007-1} observed.

\section{Historic Ties Between Cognitive Science and Computer Science}
At the dawn of the computing revolution, visionaries predicted that
computers had the potential to “augment the human intellect”
\autocite{engelbart_augmenting_1962}, a much more expansive view than
simply becoming tools to automate tedious calculations. What was meant
by this was that computing allows for new ways of dealing with
complexity, provides new metaphors for thinking about
systems\autocite{kay_history_1996}. Early work demonstrated that with
the aid of computers, children could engage with tasks and concepts
that had previously been considered “advanced”, and through
interacting with computers, children could become more aware of their
own cognitive processes \autocite{papert_mindstorms_1980}. However,
\textcite{east_patternbased_1996} note that as of 1996 little work had
been done to inform instruction from knowledge of learning. The rise
in popularity of object-oriented programming seems to dominate the
conversation for the next few years, though
\textcite{robins_learning_2003} observes there is a lack of evidence
supporting any claim that OOP makes modeling problems any easier.

%“Computing enters as a new representation system with new and useful
%metaphors for dealing with complexity, especially of systems”
%\autocite{kay_history_1996}

% “I wasn't smarter but I had a much better internal thinking tool to
% amplify my abilities.” \autocite[][p.526]{kay_history_1996} about his
% ability to quickly solve a problem using LISP.

% “the 'hacker phenomenon'” \autocite[][p.544-5]{kay_history_1996}

% Difficulty novices experience have “little to do with features of the
% programming languages used, and everything to do with the difficulties
% novices have thinking the secial way that good programmers think”
% \autocite[][p.546]{kay_history_1996}

% “However, a program is not just for the computer: A program that
% follows the rules of discourse better enables a human reader to
% reconstruct the explanation that the program writer was following in
% initially developing the mechanism.”
% \autocite[][p.857]{soloway_learning_1986}

\section{Methods}
Assuming that there would be more comprehensive research of
introductory programming in the computer science community we began
with targeted searchs of the ACM digital library for `introductory
programming'. Once we found key papers that seemed to indicate a call
to action
\autocite[e.g.][]{mccracken_multinational_2001,kolling_problem_1999},
or a survey of practices
\autocite[e.g.][]{pears_survey_2007-1,robins_learning_2003} we used
the `cited by' list to find additional relevant papers.

%% Despite hope in the 1990s that computing and programming had potential
%% for bringing more undrerepresented groups into STEM fields, it has
%% become one of the *least* diverse fields.

% \textcite{turkle_epistemological_1990} were optimistic that paradigm
% shift to object oriented programming that began in the 1990s would
% result in more epistemological diversity since at its core, OOP as a
% design philosophy is compatible with what they call a “bricolage”
% strategy towards design (e. g. the concept of message passing between
% agents invokes association with a dialog or negotiation which is how
% people following a bricolage style see themselves interacting with the
% machine). Unfortunately, based on more current literature regarding
% how OOP is taught the elements of OOP that most align with a bricolage
% style are left out while concepts of OOP that align with the
% traditional “planner” style: encapsulation, inheritance are
% emphasized.

\section{Current state of research}
The bulk of literature about teaching introductory programming (CS1)
revolves around what paradigms are “best” (e.g. procedural vs. object
oriented), which language is “best”, often times with regard to a
particular paradigm (e.g. OOP with Python vs. Java), and best teaching
practices (e.g. use of “pair programming” environments). However, the
\emph{reasons} for picking one paradigm over another tend to be about
industry demand, rather than intentionally choosing a paradigm to
support the learning objectives of the course.

\section{Overall direction}
A \citeyear{robins_learning_2003} study by
\citeauthor{robins_learning_2003} identifies four trends in research
into programming education: 1) distinguishing between novice and
expert programmers with emphasis on deficiencies of novices, 2) the
distinction between knowledge and strategies, 3) the distinction
between program comprehension and generation, and 4) a comparison of
OOP to procedural styles of programming\autocite{robins_learning_2003}.

The \textcite{mccracken_multinational_2001} study, which concluded that
regardless of nation or institution, most students do not know how to
program by the end of their introductory programming courses, sparked
much conversation and push to reform introductory programming in the
next decade. Ten years after the
\citeauthor{mccracken_multinational_2001} study,
\textcite{lister_ten_2011} reflected on the response and lemented that
while much of the criticism that the problems used to test student
knowledge were too difficult, no one had repeated the study with
problems they deemed to be more reasonable. Thus while the
\citeauthor{mccracken_multinational_2001} study validated the
experiences of many introductory programming instructors it did not
result in critical change to address the problems. Rather, what we
have seen since the 1990s is been an increasing interest in where
object oriented programming should be placed within the curriculum
with debates between objects-first vs procedural-fist but little
question of why object oriented design should be included at all other
than it continues to be a popular paradigm in industry.

Much fewer, but still some, authors have drawn on cognitive sciences
to determine how indroductory programming courses should be
taught. \textcite{muller_pattern_2005} propose a pattern oriented
approach that places emphasis on recognizing and learning to work with
algorithmic patterns over syntax and design paradigms.

There remains a desire that introductory programming courses should
teach problem-solving skills, but few papers present evidence that the
implementation of this goal is informed by cognitive science.

\citeauthor{pears_survey_2007-1} conclude their
\citeyear{pears_survey_2007-1} literature survey with the observation
that despite the volume of work exploring the implementation of
introductory programming courses “there is little systematic evidence
to support any particular approach.”\autocite{pears_survey_2007-1} This
could be due to two reasons:

\begin{enumerate}
\item There has been a lack of consensus on what the goals of an
  introductory course should be: the ability to write a complete,
  working program, the ability to answer conceptual questions about
  computing and computer science, or the ability to analyze problems
  and design solutions (problem-solving). The “appropriate” approach
  to teaching would depend on the goal.

\item Even if goals were clearly stated it is likely many different
  approaches could be used successfully if chosen intentionally to
  align with the goal and measurements of success (assessment). In
  many of the studies the metrics used to determine success suffer
  from validity issues, e.g. the validity of final grades on a final
  assessment depend tremendously on the design of that assessment and
  there is little evidence to suggest course assessments are
  intentionally designed with learning goals in mind.
\end{enumerate}

A large number of papers surveyed contain discussions of course
objectives, pedagogical tools used, and assessments given, but rarely
are these three components discussed together.

\subsection{Components of a learning
system}\label{sec:components-of-a-learning-system}

A popular way to visualize an effective learning system is as a triad
of objectives, instruction, and
assessment\textcite{felder_designing_2003}, shown in Figure
\ref{fig:learning-system}. The design of these three components is not
sequential: each informs the other.

\begin{figure}[h!]
  \centering
  \includestandalone[width=\textwidth]{felder_framework}%
  \caption{Learning system}
  \label{fig:learning-system}
\end{figure}

\subsection{Effective learning systems}\label{sec:effective-learning-systems}

Assessment results can help inform changes to course objectives and
instructional methods\autocite{felder_designing_2003}. Similarly, as
instructional methods improve, the objectives may be modified to aim
for higher levels of understanding. Being intentional about these
interactions is crucial to an effective learning system; if the
assessment addresses lower levels of learning than the objectives
specify, ``the system will be driven by backwash from testing, not by
the curriculum.''\autocite[][p. 350]{biggs_enhancing_1996}. For
example, if an objective of a programming course is that learners will
be able to transfer high level concepts from programming to other
domains, but assessment only consists of checking that programs
function conforms to data input/output specifications,
teaching/learning activities would tend to center around those that
promote correct input/output behavior and exclude activities that
might have a better chance at promoting higher level understanding and
transfer\autocite{biggs_enhancing_1996}.

\section{Case Study}
\subsection{Object Oriented Programming}
As noted, teaching OOP is a common area of interest in the literature.

It should be strongly noted that we are not advocating for OOP as a
``best'' teaching paradigm. Rather, we want to point out that from what
has been reported it seems that it is often not taught
effectively. What educators seem to miss is that OOP defines an
entirely different paradigm for problem solving than is used in
procedural programming\autocite{kolling_problem_1999}. As should not
be surprising there are some problems for which OOP provides a good
set of tools to solve in an efficient, elegant manner, namely those
with solutions taht naturally map to real-world
objects\autocite{raymond_art_2003}. Different problems may be better
solved using a procedural, or functional
paradigm\autocite{kay_history_1996,felleisen_functional_2009,crestani_experience_2010}. Ultimately,
our goal should be to help students to analyze problem structures and
make informed decisions regarding what paradigm would be best for a
particular problem. What is not clear from the literature is whether
or not this consideration is made when designing problems for a
particular course. Indeed, in some cases there is evidence that OOP
concepts are added to a course ``on top of'' older content, without
necessarily changing the problem sets (citation needed).

Alan Kay originally conceived of object oriented programming as as one
way to think about complex systems\autocite{kay_history_1996}. The
underlying metaphore of Kay's conception of OOP was that of the
biological cell: each object is designed to accomplish a specific set
of tasks where the complexity of these tasks is encapsulated within
the object. Objects interact with one another by sending and receiving
messages and systems of these objects are built to solve complex
problems\autocite{kay_history_1996}.

[section on what an ideal alignment would look like using OOP]

\section{Language choice}
Choice of language does matter: If the learning objectives of an
introductory programming course are to learn problem solving concepts
then a language with a simpler syntax is preferable to one with more
complex syntax features\autocite{koulouri_teaching_2014}. In an
attempt to address the concern that the two primary languages, C++ and
Java, used to teach OOP in introductory courses, may be too complex
for novice students, \textcite{goldwasser_teaching_2008} chooses
Python to be the language of choice for introductory programming
claiming that its relatively simple syntax allow new students to
better engage with the OOP concepts rather than getting bogged down by
details of the language.

% \section{What are “the basics”}
% Several papers discuss the relative merits of a more “modern” paradigm
% such as OOP and procedural paradigms and what they refer to as “basic
% concepts” in programming. The “basics” tend to be conditional
% structures, looping structures, and arrays. However it is never noted
% that these “concepts” are only “basic” in the context of a particular
% paradigm, in particular procedural. For example, if we decided it was
% best to teach a functional paradigm in CS1 we would be hard pressed to
% defend conditional and looping structures as “basic”, rather we would
% consider recursion, evaluation, and higher-order-functions as “basic”
% concepts.

% \section{The relative dearth of functional-first approaches}
% Given the apparent focus on formalist design practices, what
% \textcite{turkle_epistemological_1990} refer to as “the planner”
% approach, it is interesting to note the relative lack of interest in
% using functional-first strategies. The functional programming paradigm
% is highly compatible with formal and abstract design methods and is
% closely associated with mathematical problem solving
% \autocite{backus_can_1978}. The relative lower interest in functional
% programming suggests that programming paradigms are chosen more for
% practical market concerns (e.g. OOP is popular in industry, adoption
% of functional styles have been slow to adoption largely due to early
% problems with speed), rather than any conscious link between design
% strategies, objectives and paradigm.

% \section{Epistomological Pluralism}
% \textcite{vilner_fundamental_2007} argues that providing “different
% ways of approaching and solving problems” should be a goal of
% introductory programming courses.

\section{Discussion}

\section{Future Work}

There continues to be a disconnect between learning sciences and how
programming paradigms and languages are selected for introductory
courses. While there is quite a bit of exploration around connections
between abstract thinking, threshold concepts, etc. and ``success'',
“success” is not consistently defined, sometimes meaning the ability
to write programs (e.g. with specifications given), sometimes focusing
more on the ability to solve problems.

\subsection{Continued Ties to Learning Sciences}

The thread of learning science influence split into two branches, one
further exploring the pattern-based approach and one exploring
threshold concepts. While developing problem-solving skills remains a
goal of CS1 only the pattern-based approach thread seems to focus on
this aspect while the researchers exploring threshold concepts are
more concerned with overall success in computer science without
specific mention of problem-solving skills.

\subsubsection{Problem Solving}

Several studies pick up the pattern-based focus of programming to
encourage problem-solving skill development from the
1990s\autocite{muller_almost_2004,muller_pattern_2005,muller_patternoriented_2007,caspersen_instructional_2007-1}.

In one of these studies, the authors draw from three concepts of
cognitive science: cognitive load theory, cognitive apprenticeship,
and the theory of worked examples, to design an introductory
programming course\autocite{caspersen_instructional_2007-1}. It is
interesting to note that they select these theories based on a human
cognition model that resulted from computing: separate working and
long term memory where learning is the process for transferming
information from working to long-term memory. The course the authors
describe teachs OOP but they never explain why the chose OOP as the
paradigm, it is assumed that this is just a given feature of an
introductory programming course.

\subsubsection{Threshold Concepts}

Some studies explore potential threshold
topics\autocite{meyer_threshold_2005} for computer science and how
they might be useful for informing course
design\autocite{mccartney_liminal_2009,boustedt_threshold_2007}. \citeauthor{boustedt_threshold_2007}
identify object oriented programming and pointers as candidate
threshold topics based on interviews with students graduating with
computer science degrees. The two concepts and so these concepts are
in the context of computer science in general rather than introductory
programming, thus there is not a clear link between these concepts and
the goal of problem solving. In addition the authors note that OOP in
particular is quite broad and suggest further work to narrow the focus
of that threshold topic to one ore more specific aspects of OOP, such
as how objects communicate between one another.

\subsubsection{Objects-first vs. X}
Much of the literature focuses on the debate between introducing
object oriented programming in CS1 or whether introductory courses
should teach a procedural style. Several researchers argue that an
object-first approach emphasices coding constructs over other goals
such as problem-solving and design
\autocite{moritz_objectsfirst_2005,muller_guidelines_2005,reges_back_2006}. \textcite{moritz_objectsfirst_2005}
propose a “design-first” approach that gets students to analyze
problems and design solutions before writing any computer code.

\autocite{vilner_fundamental_2007} found no significant difference in
student performance on questions relating to recursion and efficiency
between groups that were given an object-first curriculum vs. groups
given a procedural-first curriculum. They did observe a significant
difference between the two groups when asked to design an abstract
data type: those who were in the objects-first group performed better
than those in the procedural-first group. However, it is notable that
the students were only tested on programming knowledge, not their
ability to analyze problems and design solutions.

\textcite{reges_back_2006} takes a ``back to basics'' approach eschewing
the popular OOP and reverts his CS1 course back to a procedural style
which allows him to put the emphasis back on problem solving.

\vspace{4\baselineskip}\vspace{-\parskip} % Creaters proper 4 blank line spacing.

\printbibliography

\end{document}