### 2016-02-03 paper.tex
another wave of edits
filling in some gaps, adding future work and conclusion

\documentclass[12pt]{article}

% sorting=none will order citations in order of appearance
% use sorting=nyt to sort by name, year, title
\usepackage[backend=biber,autolang=hyphen, bibencoding=inputenc, isbn=false, uniquename=false, sorting=none, sortcites=true, style=numeric, autocite=superscript, url=false]{biblatex} % biblatex setup
\DeclareLanguageMapping{american}{american-apa}
\appto{\bibfont}{\footnotesize} % for 10pt bibliography
%\usepackage{xunicode}
\usepackage[american]{babel}
\usepackage[babel]{csquotes}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows,shapes,fit}
\usepackage{standalone}
%\usepackage{mdframed}
\usepackage{changepage}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{mdframed}

\titleformat*{\section}{\normalsize\bfseries} % Makes section titles 12 pt font
\titleformat{\section}{\normalsize\bfseries}{}{}{}[]
\titleformat{\subsection}{\bfseries}{}{}{}[]
\titleformat{\subsubsection}{\bfseries}{}{}{}[]

%%\let\textcite=\autocite
\addbibresource{computing.bib}

\title{\large \textbf{Leveraging Historical Ties between Cognitive Science and
  Computer Science to Guide Programming Education}}

% Author info isn't included for the Annual Conference but some regional conferences might request it.
\author{}
%\author{\normalsize Author Name\\
%\normalsize email@example.com\\
%\normalsize Name of Your Department\\\
%\normalsize Your Institution Name}
\date{}

\makeatletter % This gets the margins for the title set.
\patchcmd{\@maketitle}{\begin{center}}{\begin{adjustwidth}{0.5in}{0.5in}\begin{center}}{}{}
\patchcmd{\@maketitle}{\end{center}}{\end{center}\end{adjustwidth}}{}{}
\makeatother

%% TODO: be more explicit about learning framework

%% TODO: bring in formative assessment

\begin{document}
\raggedright
\maketitle
\thispagestyle{empty}
\pagestyle{empty}

% conflict between teaching programming syntax, teaching ``how to program'', teaching high-level concepts like abstractions and how to ``do'' problem solving.
\section{Coding for everyone}
In the past few years, there has been increasing interest to get more
people, regardless of background, to learn to program. In fact,
President Obama recently made a statement calling on all children to
have the opportunity to learn about computer
science\autocite{whitehouse_computer_2016}. Sites such as code.org
promote CS education opportunities for all, citing statistics about
STEM jobs and arguing that in the 21st century knowledge about
computer science is foundational\autocite{code.org_every_2016}. Many
institutions of higher education either have programming requirements
for non-computer-science majors, or have been expanding programming to
non-majors\autocite{rich_cs1_2004,forte_motivation_2005,guzdial_design_2005}. At
the authors of this paper's institution, all first year general
engineering students are required to complete an introductory
programming module as part of their first-year engineering course, and
in general ``basic programming skills'' is a common desired outcome
across engineering
curricula\autocite{reid_classification_2014}.

This coding-for-all paradigm juxtaposed with the observation that
technology is changing rapidly, and the half-life of technical skills
is decreasing\autocite{nae_educating_2004}, it would seem that
teaching programming is a questionable use of class time: the language
and paradigm students are taught will likely not be the same they are
asked to use in their first job post-graduation. However, if we
acknowledge that learning to program is more than simply learning to
write code, and in particular if we draw on work from the early era of
computing which closely linked programming to problem solving and
cognitive skills, we can argue that regardless of language or
paradigm, learning to program may actually be a way to learn to
problem solve, organize knowledge, and conceptualize processes.

At the dawn of the computing revolution, visionaries predicted that
computers had the potential to ``augment the human intellect''
\autocite{engelbart_augmenting_1962}, a much more expansive view than
simply becoming tools to automate tedious calculations. What was meant
by this was that computing allows for new ways of dealing with
complexity, provides new metaphors for thinking about
systems\autocite{kay_history_1996}. Early work demonstrated that with
the aid of computers, children could engage with tasks and concepts
that had previously been considered ``advanced'', and through
interacting with computers, children could become more aware of their
own cognitive processes \autocite{papert_mindstorms_1980}. Thus, the
fact that specific technologies may be obsolete by the time students
graduate may be beside the point, there are lasting cognitive benefits
learning to interact with a computer.

While there are some recent studies which investigate programming
education as a means of teaching problem-solving, there is also a
large body of research that primarily focuses on teaching students the
details of how to code. We argue that in most cases these details are
less important than centering programming education around
problem-solving skills. If we make learning problem-solving skills a
goal of introductory programming courses, then all aspects of the
course, from the course objectives, to the curriculum used, to the
assessments administered, must be chosen intentionally with this goal
in mind.

In this paper we will present a brief summary of current published
work on introductory programming education. In particular we will
explore the alignment of course objectives, pedagogical strategies,
and assessments, and will offer a potential framework to help think
about both the design and evaluation of introductory programming
courses.

%“Computing enters as a new representation system with new and useful
%metaphors for dealing with complexity, especially of systems”
%\autocite{kay_history_1996}

% “I wasn't smarter but I had a much better internal thinking tool to
% amplify my abilities.” \autocite[][p.526]{kay_history_1996} about his
% ability to quickly solve a problem using LISP.

% “the 'hacker phenomenon'” \autocite[][p.544-5]{kay_history_1996}

% Difficulty novices experience have “little to do with features of the
% programming languages used, and everything to do with the difficulties
% novices have thinking the secial way that good programmers think”
% \autocite[][p.546]{kay_history_1996}

% “However, a program is not just for the computer: A program that
% follows the rules of discourse better enables a human reader to
% reconstruct the explanation that the program writer was following in
% initially developing the mechanism.”
% \autocite[][p.857]{soloway_learning_1986}

\section{Methods}
Assuming that there would be more comprehensive research of
introductory programming in the computer science community we began
with targeted searchs of the ACM digital library for `introductory
programming'. Once we found key papers that seemed to indicate a call
to action
\autocite[e.g.][]{mccracken_multinational_2001,kolling_problem_1999},
or a survey of practices
\autocite[e.g.][]{pears_survey_2007-1,robins_learning_2003} we used
the `cited by' list to find additional relevant papers.

Because the literature search of introductory programming seemed to be
interest in why and how programming was taught, we needed a documented
framework to consistently evaluate and starting from
\citeauthor{jamieson_creating_2009}'s
(\citeyear{jamieson_creating_2009}) to ground educational research
activities in how people learn we identified the concept of
\emph{constructive alignment} as a promising one. Following the
references in \citeauthor{jamieson_creating_2009}'s report as well as
conducting targeted literature searches for ``constructive alignment''
we were able to find descriptions of this
framework\autocite{jamieson_creating_2009,pellegrino_rethinking_2006,felder_designing_2003,biggs_enhancing_1996}.
%% TODO: mention that we chose the framework after analyzing the above
%% citations, i.e. it wasn't a direct pull from jamieson

%% Despite hope in the 1990s that computing and programming had potential
%% for bringing more undrerepresented groups into STEM fields, it has
%% become one of the *least* diverse fields.

% \textcite{turkle_epistemological_1990} were optimistic that paradigm
% shift to object oriented programming that began in the 1990s would
% result in more epistemological diversity since at its core, OOP as a
% design philosophy is compatible with what they call a “bricolage”
% strategy towards design (e. g. the concept of message passing between
% agents invokes association with a dialog or negotiation which is how
% people following a bricolage style see themselves interacting with the
% machine). Unfortunately, based on more current literature regarding
% how OOP is taught the elements of OOP that most align with a bricolage
% style are left out while concepts of OOP that align with the
% traditional “planner” style: encapsulation, inheritance are
% emphasized.

\section{Current state of research}
The bulk of literature about teaching introductory programming (CS1)
revolves around what paradigms are ``best'' (e.g. procedural
vs. object oriented
\autocite{bailie_objects_2003,moritz_objectsfirst_2005,reges_back_2006}),
which language is ``best'', often times with regard to a particular
paradigm (e.g. OOP with Python
vs. Java\autocite{goldwasser_teaching_2008,koulouri_teaching_2014}),
and best teaching practices (e.g. use of “pair programming”
environments\autocite{williams_support_2001,mcdowell_effects_2002,thomas_code_2003,chong_social_2007,salleh_empirical_2011}). However,
the \emph{reasons} for picking one paradigm over another tend to be
about industry demand, rather than intentionally choosing a paradigm
to support the learning objectives of the course. This behavior,
making a popular practice part of a learning environment without
critically analyzing whether or not is a good fit is by no means
exclusive to programming education. For example, flipped
classrooms\autocite{bishop_flipped_2013}, in which students are asked
to watch recorded lectures at home and use class time for working on
activities, has recently been a hot topic in higher education and
there have been a continuous number of examples of well-intentioned
practitioners flipping their classroom without adjusting course
objectives, content and activities to work under the flipped-classroom
model\autocite{desantis_students_2015}. Likewise, anyone can teach
elements of object oriented programming in their introductory
programming courses, but unless critical aspects of the course are
adjusted: background reading, rethinking what the ``fundamental
concepts'' are, the types of problems assigned, and methods of
assessment, the incorporation of OOP will likely create confusion for
students and frustration for instructor\autocite{kolling_problem_1999}.

%% TODO: more directed title
%% TODO: possibly use subheadings to ``tweet'' the major takeaways from each section,
%% this could then be connected at the end to bullet points of suggestions
\section{Trends in programming education research}
A \citeyear{robins_learning_2003} study by
\citeauthor{robins_learning_2003} identifies four trends in research
of programming education: 1) distinguishing between novice and expert
programmers with emphasis on deficiencies of novices, 2) the
distinction between knowledge and strategies, 3) the distinction
between program comprehension and generation, and 4) a comparison of
OOP to procedural styles of
programming\autocite{robins_learning_2003}. They found that the
biggest challenge for novice programmers was not understanding the
basic concepts, but knowing how to apply them. This finding is
consistent understanding of how we
learn\autocite{bransford_how_2000,ambrose_chapter_2010}, and with
studies of expert/novice behavior in other domains, such as
statics\autocite{litzinger_cognitive_2010}. In both cases the
suggestion is that helping students learn strategies that work is
crucial to helping them become effective novices, i.e. devote
instructional time to problem-solving strategies and how to apply
them.

A \citeyear{mccracken_multinational_2001} study by
\citeauthor{mccracken_multinational_2001}\autocite{mccracken_multinational_2001},
which concluded that regardless of nation or institution, most
students do not know how to program by the end of their introductory
programming courses, sparked much conversation and push to reform
introductory programming in the next decade. This phenomenon is by no
means exclusive to introductory programming: disconnect between what
we think we are teaching and what students learn has been observed
across
engineering\autocite{streveler_learning_2008,flynn_engineering_2014}. Possible
ways of addressing this disconnect include recognizing differences in
student's prior knowledge, helping students see connections between
abstract concepts, and relating difficult concepts to ones students
already know\autocite{streveler_learning_2008,slotta_helping_2006}.
%% TODO: connect mccracken to engineering ed about missconceptions
%% (CHEER chapter and Streveler paper)

While the \citeauthor{mccracken_multinational_2001} study validated the
experiences of many introductory programming instructors it did not
result in critical change to address the problems. Rather, what we
have seen since the 1990s is been an increasing interest in where
object oriented programming should be placed within the curriculum
with debates between objects-first vs procedural-fist but little
question of why object oriented design should be included at all other
than it continues to be a popular paradigm in industry.

There remains a desire that introductory programming courses should
teach problem-solving skills, though only a few papers present
evidence that the implementation of this goal is informed by cognitive
science\autocite{kramarski_cognitivemetacognitive_1997,reed_incorporating_1998,muller_supporting_2008}. One
thread in particular has received regular attention:
\citeauthor{east_patternbased_1996} propose a pattern-based approach
that places emphasis on recognizing and learning to work with
algorithmic patterns over syntax and design
paradigms\autocite{east_patternbased_1996}. This approach closely
aligns with understanding of differences in expert/novice behavior by
helping students identify patterns of solutions and learning how and
when to apply them to particular
problems\autocite{east_patternbased_1996,muller_pattern_2005,wallingford_first_1996}.

\citeauthor{pears_survey_2007-1} conclude their
\citeyear{pears_survey_2007-1} literature survey with the observation
that despite the volume of work exploring the implementation of
introductory programming courses “there is little systematic evidence
to support any particular approach.”\autocite{pears_survey_2007-1}
We think this could be due to two reasons:

\begin{enumerate}
\item There has been a lack of consensus on what the goals of an
  introductory course should be: the ability to write a complete,
  working program, the ability to answer conceptual questions about
  computing and computer science, or the ability to analyze problems
  and design solutions (problem-solving). The “appropriate” approach
  to teaching would depend on the goal.

\item Even if goals were clearly stated it is likely that many
  different approaches could be used successful if chosen
  intentionally to align with the goal and measurements of success
  (assessment). In many of the studies it is not immediately clear if
  the measures of success used are appropriate for the stated
  objectives. That is to say, the assessments used might not always be
  measuring all aspects of the desired learning objectives.
\end{enumerate}

Finally, in \citeyear{koulouri_teaching_2014},
\citeauthor{koulouri_teaching_2014} conducted a quantitative analysis
of different approaches to teaching introductory programming by
varying three aspects: the language used, providing formative
feedback, and introducing explicit problem solving
training\autocite{koulouri_teaching_2014}. They found that choosing a
language with a simpler syntax (Python instead of Java), and
introducing explicit problem-solving training improved student
learning, while providing more formative assessment did not seem to
help. They speculate that the reason for the latter finding could be
that students do not spend time looking at formative feedback, or do
not understand what is given.

We wish to build upon the past reviews of the field by suggesting an
approach to designing learning systems that may help to consistently
address some of the shortcomings observed, and bring understanding to
the successes. A large number of papers surveyed contain discussions
of course objectives, pedagogical tools used, and assessments given,
but rarely are these three components discussed together. We know from
the engineering education literature that an effective learning system
depends on the alignment of these three
components\autocite{jamieson_creating_2009,pellegrino_rethinking_2006}.

\subsection{Components of a learning
system}\label{sec:components-of-a-learning-system}

A popular way to visualize an effective learning system is as a triad
of objectives, instruction, and
assessment\autocite{felder_designing_2003}, shown in Figure
\ref{fig:learning-system}. The design of these three components is not
sequential: each informs the other.

\begin{figure}[h!]
  \centering
  \includestandalone[width=\textwidth]{felder_framework}%
  \caption{Learning system as derived from
    \citeauthor{felder_designing_2003}\autocite{felder_designing_2003}}
  %%TODO: as derived from original figure
  \label{fig:learning-system}
\end{figure}

\subsection{Effective learning systems}\label{sec:effective-learning-systems}

Assessment results can help inform changes to course objectives and
instructional methods\autocite{felder_designing_2003}. Similarly, as
instructional methods improve, the objectives may be modified to aim
for higher levels of understanding. Being intentional about these
interactions is crucial to an effective learning system; if the
assessment addresses lower levels of learning than the objectives
specify, ``the system will be driven by backwash from testing, not by
the curriculum.''\autocite[][p. 350]{biggs_enhancing_1996}. For
example, if an objective of a programming course is that learners will
develop problem solving skills, but assessment only consist of test of
programming concepts and syntax, teaching/learning activities would
tend to center around those that promote understanding of these
concepts and exclude activities that might have a better chance at
promoting general problem solving\autocite{biggs_enhancing_1996}.

\section{A hypothetical case study}
As noted, teaching OOP is a common area of interest in the
literature. While \citeauthor{robins_learning_2003} note there is a
lack of evidence supporting any claim that OOP makes modeling problems
any easier\autocite{robins_learning_2003}, we will incorporate it into
a hypothetical introductory programming course to explore how its
inclusion would affect the constructive alignment between course
objectives, instructional methods, and assessment.

It should be strongly noted that we are not advocating for OOP as a
``best'' teaching paradigm. Rather, we want to point out that from
what has been reported teaching OOP effectively can be
challenging. What many implementations seem to miss is that OOP
defines an entirely different paradigm for problem solving than is
used in procedural programming\autocite{kolling_problem_1999}. As
should not be surprising there are some problems for which OOP
provides a good set of tools to solve in an efficient, elegant manner,
namely those with solutions that naturally map to real-world
objects\autocite{raymond_art_2003}. Different problems may be better
solved using a procedural, or functional
paradigm\autocite{kay_history_1996,felleisen_functional_2009,crestani_experience_2010}. Ultimately,
our goal should be to help students analyze problem structures and
make informed decisions regarding what paradigm would be best for a
particular problem. What is not clear from the literature is whether
or not this consideration is made when designing problems for a
particular course. Indeed, in some cases there is evidence that OOP
concepts are added to a course ``on top of'' older content, without
necessarily changing the problem
sets\autocite{kolling_problem_1999,vilner_fundamental_2007}.

\subsection{Object oriented programming}

Alan Kay originally conceived of object oriented programming as
\emph{one} way to think about complex
systems\autocite{kay_history_1996}. The underlying metaphore of Kay's
conception of OOP was that of the biological cell: each object is
designed to accomplish a specific set of tasks where the complexity of
these tasks is encapsulated within the object. Objects interact with
one another by sending and receiving messages, and systems of these
objects are assembled to solve complex
problems\autocite{kay_history_1996}.

%%TODO: link to knowledge/skills interaction literature
%%TODO: bring in knowledge organization, focus on concepts vs. procedures

In several reports of attempts to incorporate OOP into introductory
programming courses, practitioners worried that time devoted to OOP
concepts would necessarily reduce time spent on the ``basics of
programming'', usually defined as conditional statements, loops, and
in some cases pointers. This concern is misplaced, under an OOP
paradigm the ``basics of programming'' are not the same as when
working from a procedural paradigm. OOP defines a completely different
set of tools for problem solving, and it is these tools and concepts
that become the
``basics''\autocite{kolling_problem_1999,kay_history_1996}.

For instance, if taking a patterns-based approach to teaching
programming one could use either a procedural paradigm or an OOP
paradigm, but the fundamental patterns would be quite different
depending on the paradigm\autocite{wallingford_first_1996}.

Ultimately, if our goal is to teach problem solving in general, we
would be doing a grave disservice to our students to leave with the
impression that there is only one paradigm, whether it be OOP,
procedural, functional, or something not yet invented. As previously
mentioned, OOP and procedural paradigms are comprised of fundamentally
different concepts for problem solving, and so again is the functional
paradigm. Unsurprisingly, there is no paradigm that is the ``best''
fit for every problem. Leaving students with the believe that only one
paradigm exists makes it very difficult for them to learn how to
intentionally select and apply the ``best'' paradigm to solve a
particular problem. \citeauthor{kolling_problem_1999} argues that
performing a paradigm shift from procedural style to OOP is difficult,
and so concludes that if OOP is a desired outcome then it should be
taught \emph{first} since the reverse shift, from OOP to procedural,
is conceptually easier.

\subsection{Language choice}
Choice of language does matter: If the learning objectives of an
introductory programming course are to learn problem solving concepts
then a language with a simpler syntax is preferable to one with more
complex syntax features\autocite{koulouri_teaching_2014}. The
explanation for this can be found in cognitive load theory (CLT),
which posits that learners have a finite capacity for their short-term
memory, and for learning to occur, all aspects of the task must fit
into this finite space. While some contributors to cognitive load are
necessary for the task at hand, others are
not\autocite{paas_cognitive_2003}. If, for example, the goal is to
teach problem-solving techniques, there will be cognitive load
associated with thinking about different strategies one could
use. Learning the syntax rules of a programming language adds to the
cognitive load, but it does not aid in the process of learning
problem-solving techniques. Thus, it is in the educator's best
interest to minimize cognitive load associated with features that do
not directly help with the learning objectives of the course and
select a teaching language with easy-to-learn syntax.

In an attempt to address the concern that the two primary languages,
C++ and Java, used to teach OOP in introductory courses, may be too
complex for novice students, \citeauthor{goldwasser_teaching_2008}
select Python as a good choice for introductory programming claiming
that its relatively simple syntax allow new students to better engage
with the OOP concepts rather than getting bogged down by details of
the language\autocite{goldwasser_teaching_2008}.

\subsection{Problem Solving}
Developing problem solving was widely regarded as a common objective
for introductory programming courses. Achieving this objective is more
likely if the course begins with a focus on problem solving
strategies, before any programming is
introduced\autocite{koulouri_teaching_2014}. Separating
problem-solving instruction in this way has a number of benefits
grounded in cognitive science research: first, it serves to activate
prior knowledge\autocite{ambrose_how_2010}. With problem-solving
strategies fresh in their memory, learners will more easily apply them
when they begin programming. Second, it helps manage cognitive
load. Where as in the previous example of language choice we reduced
cognitive load by choosing a language with a simpler syntax, here we
separate two tasks that we know individually demand a high level of
cognitive load for novices: learning problem-solving strategies, and
learning programming techniques.

\section{Choice of problems to assign}
As previously noted, OOP provides a set of tools that naturally lend
themselves to solving certain types of problems but not others. An
educator teaching OOP should carefully select problems and challenges
that make sense to be solved with an OOP approach. Of course,
providing illustrative examples of problems that are \emph{not} best
solved with an OOP approach would be useful provided there is also
time for discussion of \emph{why} OOP might not be suited for certain
problems and brief mention of alternative approaches.

\section{Assessment}
Often times assessment becomes the weak point of a learning
system. Designing good assessment is
hard\autocite{lister_blooming_2000,traynor_synthesis_2005}, and often good assessment can
require significant time on the part of the instructor. Because of
this, assessment is often the weak point, lowering the effectiveness
of the entire system.

If our course objective is to teach learners to solve problems with
programming then an ideal assessment would be to give students a
problem to solve, ask them to write a program to solve it as well as
reflect upon the patterns they used in their solution. Asking for
reflection prompts students to engage in meta-cognition, becoming
aware of their own thinking process, which helps them become more
effective
learners\autocite{ambrose_how_2010,chi_selfexplanations_1989}. Of
course evaluating this type of assessment is much more time consuming
than a multiple-choice exam that can be automatically scored, or even
than evaluating that the finished program runs correctly, which can be
done automatically as well\autocite{cheang_automated_2003}.

\section{Considerations and Limitations}
In the above hypothetical example we had the luxury of complete
freedom in designing all aspects of an introductory programming
curriculum. We recognize this is almost always never the case in
practice. However, we argue that following a framework such as
constructive alignment as close as possible will help. Thinking about
the framework will also encourage critical engagement with the
constraints imposed by the institution. For example, there has been
much interest in assessment tools used for introductory programming
courses\autocite{lister_blooming_2000,lister_objectives_2001,sheard_exploring_2011},
and in particular automated
assessment\autocite{cheang_automated_2003,traynor_synthesis_2005,ihantola_review_2010}. In
many cases the push towards automated assessment techniques is a
response to increasing class sizes that instructors typically have
little control over. In the event that a large class size dictates
that some form of automated assessment be used, following the
constructive alignment framework would constrain the types of course
objectives and instruction that were feasible. For instance, if the
primary course objective was to teach conceptual understanding of
computer science, then a well-designed multiple-choice assessment
could be used
effectively\autocite{lister_blooming_2000,lister_objectives_2001}.

\section{Discussion and future work}

This literature review was by no means comprehensive, but we hope that
it touched upon some of the notable areas in which educators have been
experiencing challenges and successes in teaching introductory
programming. We have pointed out a number of possible connections
between computer science and cognitive science that may be worth
deeper investigation. While we have seen how cognitive load theory can
help reason about course design, there are also potentially
interesting connects between cognitive load theory and object oriented
design: objects are designed to encapsulate complexity to better
manage it. Additionally, further exploring concepts of knowledge
organization and differences between novices and exports may be a
fruitful area to investigate the use of teaching programming to aid in
analogical reasoning and abstract
thinking\autocite{bennedssen_abstraction_2008}.

\section{Conclusion}

Recognizing that introductory programming is a common component of
engineering curriculum, and that there is a national push to introduce
it to an even larger population, it is important to think carefully
about how we design programming education to impart the cognitive
skills that will generally benefit learners even after the specific
technologies they learn are no longer relevant. We strongly advocate
for the use a framework such as constructive alignment, described in
this paper, to inform the choices for course objectives, instructional
techniques, and assessments. Using such a framework well help identify
components of a course that may be contributing unnecessarily to
cognitive load, or reasons why student outcomes might not match
expectations assumed from course objectives.

% \section{What are “the basics”}
% Several papers discuss the relative merits of a more “modern” paradigm
% such as OOP and procedural paradigms and what they refer to as “basic
% concepts” in programming. The “basics” tend to be conditional
% structures, looping structures, and arrays. However it is never noted
% that these “concepts” are only “basic” in the context of a particular
% paradigm, in particular procedural. For example, if we decided it was
% best to teach a functional paradigm in CS1 we would be hard pressed to
% defend conditional and looping structures as “basic”, rather we would
% consider recursion, evaluation, and higher-order-functions as “basic”
% concepts.

% \section{The relative dearth of functional-first approaches}
% Given the apparent focus on formalist design practices, what
% \textcite{turkle_epistemological_1990} refer to as “the planner”
% approach, it is interesting to note the relative lack of interest in
% using functional-first strategies. The functional programming paradigm
% is highly compatible with formal and abstract design methods and is
% closely associated with mathematical problem solving
% \autocite{backus_can_1978}. The relative lower interest in functional
% programming suggests that programming paradigms are chosen more for
% practical market concerns (e.g. OOP is popular in industry, adoption
% of functional styles have been slow to adoption largely due to early
% problems with speed), rather than any conscious link between design
% strategies, objectives and paradigm.

% \section{Epistomological Pluralism}
% \textcite{vilner_fundamental_2007} argues that providing “different
% ways of approaching and solving problems” should be a goal of
% introductory programming courses.

% \section{Future Work}

% There continues to be a disconnect between learning sciences and how
% programming paradigms and languages are selected for introductory
% courses. While there is quite a bit of exploration around connections
% between abstract thinking, threshold concepts, etc. and ``success'',
% “success” is not consistently defined, sometimes meaning the ability
% to write programs (e.g. with specifications given), sometimes focusing
% more on the ability to solve problems.

% \subsection{Continued Ties to Learning Sciences}

% The thread of learning science influence split into two branches, one
% further exploring the pattern-based approach and one exploring
% threshold concepts. While developing problem-solving skills remains a
% goal of CS1 only the pattern-based approach thread seems to focus on
% this aspect while the researchers exploring threshold concepts are
% more concerned with overall success in computer science without
% specific mention of problem-solving skills.

% \subsubsection{Problem Solving}

% Several studies pick up the pattern-based focus of programming to
% encourage problem-solving skill development from the
% 1990s\autocite{muller_almost_2004,muller_pattern_2005,muller_patternoriented_2007,caspersen_instructional_2007-1}.

% In one of these studies, the authors draw from three concepts of
% cognitive science: cognitive load theory, cognitive apprenticeship,
% and the theory of worked examples, to design an introductory
% programming course\autocite{caspersen_instructional_2007-1}. It is
% interesting to note that they select these theories based on a human
% cognition model that resulted from computing: separate working and
% long term memory where learning is the process for transforming
% information from working to long-term memory. The course the authors
% describe teachs OOP but they never explain why the chose OOP as the
% paradigm, it is assumed that this is just a given feature of an
% introductory programming course.

% %%TODO: consider focusing on this and dropping threshold concepts and others. Potentally bring in pattern-based pedagogy

% \subsubsection{Threshold Concepts}

% Some studies explore potential threshold
% topics\autocite{meyer_threshold_2005} for computer science and how
% they might be useful for informing course
% design\autocite{mccartney_liminal_2009,boustedt_threshold_2007}. \citeauthor{boustedt_threshold_2007}
% identify object oriented programming and pointers as candidate
% threshold topics based on interviews with students graduating with
% computer science degrees. The two concepts and so these concepts are
% in the context of computer science in general rather than introductory
% programming, thus there is not a clear link between these concepts and
% the goal of problem solving. In addition the authors note that OOP in
% particular is quite broad and suggest further work to narrow the focus
% of that threshold topic to one ore more specific aspects of OOP, such
% as how objects communicate between one another.

% \subsubsection{Objects-first vs. X}
% Much of the literature focuses on the debate between introducing
% object oriented programming in CS1 or whether introductory courses
% should teach a procedural style. Several researchers argue that an
% object-first approach emphasices coding constructs over other goals
% such as problem-solving and design
% \autocite{moritz_objectsfirst_2005,muller_guidelines_2005,reges_back_2006}. \textcite{moritz_objectsfirst_2005}
% propose a “design-first” approach that gets students to analyze
% problems and design solutions before writing any computer code.

% \autocite{vilner_fundamental_2007} found no significant difference in
% student performance on questions relating to recursion and efficiency
% between groups that were given an object-first curriculum vs. groups
% given a procedural-first curriculum. They did observe a significant
% difference between the two groups when asked to design an abstract
% data type: those who were in the objects-first group performed better
% than those in the procedural-first group. However, it is notable that
% the students were only tested on programming knowledge, not their
% ability to analyze problems and design solutions.

% \textcite{reges_back_2006} takes a ``back to basics'' approach eschewing
% the popular OOP and reverts his CS1 course back to a procedural style
% which allows him to put the emphasis back on problem solving.

\vspace{4\baselineskip}\vspace{-\parskip} % Creaters proper 4 blank line spacing.

\printbibliography

\end{document}